The Galactic Delay-Time Distribution of Type Ia Supernovae:\\ A Chemical Evolution Perspective

### 2023-07-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@

\subsubsection{From Observations}

\citet{Maoz2014-Review}

@@

Another commonly assumed functional form is an exponential. \citet{Schonrich2009-RadialMixing} take their DTD to be an exponential with a 1.5 Gyr timescale, though this isn't strictly observationally motivated.

\citet{Stolger2020-ExponentialDTD} perform maximum likelihood estimations based on the cosmic star formation history and the star formation histories of individual galaxies. In both cases their best fit functions were approximately exponential, though they started with a skew-normal function so could not have ended up with a power law even if they wanted to. They don't actually report the timescale of this exponential, but based on the plots I'd say it's about 1.5 - 2 (1.7?) Gyr. Their best fit parameters make no sense to me as they don't report their units.

% \begin{deluxetable}{lCl}
% \tablecaption{Sources of the various forms of delay time distribution.}
% \tablehead{
% \colhead{Name} & \colhead{Form} & \colhead{Source}
% }
% \startdata
% Field Galaxy Power-Law      & t^{-1.1}              & \citet{Maoz2017-CosmicDTD} \\   
% Galaxy Cluster Power-Law    & t^{-1.4}              & \citet{Maoz2017-CosmicDTD} \\
% Short Exponential           & e^{-t/1.5~\rm{Gyr}}  & \citet{Schonrich2009-RadialMixing} \\
% \enddata
% \end{deluxetable}

\subsubsection{Theoretical Models}

\citet{Matteucci1986-SupernovaEnrichment}

\citet{Greggio2005-AnalyticalRates}

\citet{Rajamuthukumar2022-TripleEvolution}

\citet{Schonrich2009-RadialMixing,Weinberg2017-ChemicalEquilibrium} assume an exponential with $\tau=1.5$ Gyr

@@

\subsubsection{Our Models}

We explore five different classes of DTD: a power-law, a broken power-law with an initial flat plateau, an exponential, an exponential with a prompt Gaussian component, with one or two different parameterizations for each. Figure \ref{fig:dtds} presents these DTDs as a function of time since star formation.

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$ and normalization $C_\tau$:
\begin{equation}
    R(t) = \frac{1}{\tau} e^{-t/\tau}
    \label{eq:exponential-dtd}
\end{equation}
We investigate the case where $\tau=1.5$ Gyr, which has been used in previous studies \citep[e.g.,][]{Schonrich2009-RadialMixing,Weinberg2017-ChemicalEquilibrium} and is similar to the analytical single-degenerate DTD from \citet{Greggio2005-AnalyticalRates}. We also investigate a longer timescale of $\tau=3$ Gyr.

@@

An implementation of these DTDs in a chemical evolution model requires numerical integration (see their Equations 14, 33, and 34, respectively, for the exact functional forms), so to save on computation time we approximate the DTD for the DD prescriptions as a broken power-law with an exponential rise and tail:

\begin{equation}
    f_{\rm{Ia}}^{\rm{DD}}(t) \propto
    \begin{cases}
        t^{\alpha_1} (1 - A_{\rm{rise}} e^{-t/\tau_{\rm{rise}}}) & t \le 1~\rm{Gyr} \\
        B t^{\alpha_2} (1 + A_{\rm{tail}} e^{-(t - 1~\rm{Gyr})/\tau_{\rm{tail}}}) & t > 1~\rm{Gyr}
    \end{cases}
    \label{eq:greggio05-approx}
\end{equation}

\noindent with the normalization $B$ of the second part set such that

\begin{equation}
    B = \frac{1}{1 + A_{\rm{tail}}} t^{(\alpha_1-\alpha_2)} (1 - A_{\rm{rise}} e^{-(1~\rm{Gyr})/\tau_{\rm{rise}}}).
\end{equation}

\noindent We use a non-linear least squares method to fit Equation \ref{eq:greggio05-approx} to the analytical DTDs for DD progenitors. We obtain best-fit parameters $\alpha_1=-0.19$, $A_{\rm{rise}}=1.7$, $\tau_{\rm{rise}}=0.084$ Gyr, $\alpha_2=-0.91$, $A_{\rm{tail}}=0.85$, and $\tau_{\rm{tail}}=0.088$ Gyr for the ``WIDE'' scheme, and $\alpha_1=-0.53$, $A_{\rm{rise}}=3.0$, $\tau_{\rm{rise}}=0.042$ Gyr, $\alpha_2=-1.1$, $A_{\rm{tail}}=0.62$, and $\tau_{\rm{tail}}=0.11$ Gyr for the ``CLOSE'' scheme.\footnote{
Our implementation of these best-fit approximations, as well as the original SD and DD formulations, is available in the GitHub repository for this paper: \GitHubURL
} These best-fit approximations are represented by the dashed curves in Figure \ref{fig:dtd-greggio05} \citep[see also Figure 8 in][]{Greggio2005-AnalyticalRates}. The difference between the true analytical DTDs and our best-fit approximations is small enough as to be irrelevant for our purposes, so for the rest of this section we rely on the best-fit approximations for faster simulations.

>>>>

We explore five different functional forms for the SN Ia DTD: a power-law, a broken power-law with an initial flat plateau, an exponential, an exponential with a prompt Gaussian component, with one or two different parameterizations for each. Figure \ref{fig:dtds} presents these DTDs as a function of time since star formation.

@@

% \citet{Stolger2020-ExponentialDTD} perform maximum likelihood estimations based on the cosmic star formation history and the star formation histories of individual galaxies. In both cases their best fit functions were approximately exponential, though they started with a skew-normal function so could not have ended up with a power law even if they wanted to. They don't actually report the timescale of this exponential, but based on the plots I'd say it's about 1.5 - 2 (1.7?) Gyr. Their best fit parameters make no sense to me as they don't report their units.

\begin{deluxetable*}{lll}
\tablecaption{Summary of SN Ia DTD models explored in this paper.}
\tablehead{
\colhead{Model} & \colhead{Parameters} & \colhead{Similar to}
}
\startdata
Power-law   & $\alpha=-1.1$     & \citet[][field]{Maoz2017-CosmicDTD}; 
                                  \citet{Wiseman2021-DESRates}              \\
            & $\alpha=-1.4$     & \citet[][cluster]{Maoz2017-CosmicDTD}; 
                                  \citet{Heringer2019-FieldGalaxyDTD}       \\
Exponential & $\tau=1.5$ Gyr    & \citet[][SD]{Greggio2005-AnalyticalRates};
                                  \citet{Schonrich2009-RadialMixing,
                                         Weinberg2017-ChemicalEquilibrium}  \\
            & $\tau=3.0$ Gyr    & \\
Plateau     & $W=0.3$ Gyr, $\alpha=-1.1$    & \citet[][CLOSE DD]{Greggio2005-AnalyticalRates} \\
            & $W=1.0$ Gyr, $\alpha=-1.1$    & \citet[][WIDE DD]{Greggio2005-AnalyticalRates} \\
Prompt      & $t_{\rm max}=0.05$ Gyr, $\sigma=0.015$ Gyr, $\tau=3.0$ Gyr & \citet{Mannucci2006-TwoPopulations} \\
Triple      & & \citet{Rajamuthukumar2022-TripleEvolution}
\enddata
\end{deluxetable*}

@@

@@

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$ and normalization $C_\tau$:
\begin{equation}
    R(t) = \frac{1}{\tau} e^{-t/\tau}
    \label{eq:exponential-dtd}
\end{equation}
We investigate the case where $\tau=1.5$ Gyr, which has been used in previous studies \citep[e.g.,][]{Schonrich2009-RadialMixing,Weinberg2017-ChemicalEquilibrium} and is similar to the analytical single-degenerate DTD from \citet{Greggio2005-AnalyticalRates}. We also investigate a longer timescale of $\tau=3$ Gyr. \citet{Matteucci1986-SupernovaEnrichment}, \citet{Stolger2020-ExponentialDTD}

@@
